### 2015-04-07 src/misc/s2_discussion.tex
Updated discussion, paper 4

Based on our updating and self-motion perception studies -> is it possible to make predictions for other studies?
What would we predict van Pelt would do for instance?
+
Could check Anna/Alexandra results again -> updating + self-motion effect.

>>>>

Navigating through the environment evokes complex changes of visual, auditory, vestibular, tactile and
motor inputs to the brain. Yet, despite this motion, we perceive the world as a stable reality, maintain an integrated sense of where we are, how we are oriented, and are able to track and act rather effortlessly upon surrounding objects.  How the ability comes about it one of the key questions in neuroscience and also studied in this research project. The objective was to build computational models and perform thorough psychometric testing in order to examine the physical and biological contrast that are placed on the interaction between the vestibular and other sensory systems for spatial orientation and self-motion perception. This resulted in the following contributions to the field:

\begin{enumerate}
\item Statistical optimality can account for the way body somatosensory, neck proprioceptive and vestibular signals are integrated in spatial orientation perception. While these sensors cannot be examined in isolation, by using optimality as a starting point, their noise properties can be determined and linked to clinical deficits that are seen in particular patient groups.
\item Oculomotor signals influence self-motion perception, even in the absence of optic flow or other visual stimulation, and even when in conflict with vestibular information. 
\item This oculomotor signals should be regarded as rudimentary cue to self-motion perception; it is not veridically scaled by fixation distance in the perception of body translation. 
\item Self-motion signals interact with the dynamic perception of external world.  Errors that arise in this process suggest the use of a gaze-centered reference frames in the underlying computations. 
\end{enumerate}

In the following sections, we will provide a detailed summary of each result.

\section{Multisensory processing in spatial orientation}
Many studies have shown that the brain tends to combine noisy sensory signals in a statistically optimal way. This is normally done by showing that the noise levels of each separate modality can predict performance in combined conditions. In spatial orientation it is difficult to measure sensory signals in isolation as multisensory integration always takes place. In Chapter 2 we therefore used a statistically optimal integration model as a starting point, and attempted to account for the perceptual differences found when probing head-in-space versus body-in-space orientation. Using a psychometric approach, we tested both the perception of body tilt (subjective body tilt; SBT) as well as the perception of visual vertical (subjective visual vertical; SVV) in seven participants. Our model was able to explain the results from both tests. Because one of sensory modalities used, the noise in the neck proprioceptors, can be measured in relative isolation, we used it to independently confirm the predicted head-on-body noise found by the model. We further validated our model by showing that predictions made by our model are consistent with previously published deficits in vestibular and somatosensory patients. We conclude that Bayesian computations can account for the typical differences in spatial orientation judgments associated with different task requirements.

\section{Eye movements influence self-motion perception}
Eye movement typically accompany self-motion in order to minimize retinal slip and maximize dynamic visual acuity. In Chapter 3 we investigate whether these eye movements also have a reversed role, serving as a cue for self-motion perception. To address this question, we asked participants to compare perceived translation distance in two successive, passive, lateral whole-body translations. Eye movements during these translations were either world-stationary or body-stationary. Results show that translations were perceived shorter with body-fixed gaze compared to world-fixed gaze, indicating that eye movements indeed influence self-motion perception. Using a linear model, we estimated the relative contribution of the vestibular versus the eye movement signal, and found that eye movement signals contribute approximately 25 percent of the perceived motion. We independently validated the model by successfully predicting the effects of eye movements on self-motion perception during trails in which the eye movements were unconstrained. This further shows that eye movement signals influence self-motion perception, even in the absence of visual stimulation, and even when oculomotor and vestibular estimates are in conflict, e.g. during body-fixed gaze. We hypothesize that adverse consequences of this seemingly inflexible arrangement are minimal under natural conditions because eye movements and self-motion are highly correlated, and because eye movements are most often accompanied by veridical optic flow cues to self-motion.

\section{Partial compensation for fixation depth in self-motion perception}
The amplitude of these eye movements depend on the depth of fixation, when fixating  faraway they are smaller than when fixating nearby in the world. If these eye movements are used to augment self-motion perception, the effect of fixation depth has to be taken into account in order for self-motion perception to be veridical. In Chapter 4 we investigate whether the brain indeed compensates and takes fixation depth into account. Participants had to judge self-motion during different eye movement conditions in the absence of full-field optic flow. In a 2-AFC task, similar to the one used in Chapter 3, participants indicated whether the second of two successive passive lateral whole-body translations was longer or shorter than the first. During each translation, participants fixated either a nearby or far away target, which was either body- or world-stationary. Results show that the perceived translations were shorter for nearby world-fixed gaze compared to far away world-fixed gaze, indicating that eye movements are not properly scaled in self-motion perception. Together with the observation that self-motion perception is not affected by the depth of a body stationary fixation target, we conclude that eye movements are merely a rudimentary cue to self-motion, with a compensation for fixation depth that is partial at best.

\section{Gaze-dependent effects in spatial updating}
The brain uses these self-motion signals to update ego-centric spatial representations of the environment. In Chapter 5 we investigate how, and to which extent, the brain integrates these various self-motion signals for the spatial update. Participants were oscillated sideways while keeping gaze fixed on a stationary target. When the motion direction changed, a reference target was shown either in front or behind the fixation point. Half a cycle later, at the next reversal, we tested updating of this reference location by asking participants to judge whether a briefly flashed probe was shown to the left or right of the memorized target. Results show that both the direction and magnitude of the bias in spatial updating depends on the location of the object being updated with respect to gaze, implying that a gaze-centered reference frame is involved. We further show that these biases can be caused by an underestimated of translation amplitude, a bias of visually perceived objects towards the fovea, or by a combination of both.

\section{Concluding remarks}

To be done.

%Pieter's suggestions: 10-line summary, why optimality?, questions that remain, where do we go in the future (neurophysiology, neurostimulation, modeling, ...).

%[Could it be that the fixation target is a world-fixed reference? Add something about that here or later.]

%%%%%%%%%%%%%
% Integrated model

%Could check Anna/Alexandra results again -> updating + self-motion effect.
%Reaching poster I've made for SfN

%%%%%%%%%%%%%%
% Relation to other studies

%Based on our updating and self-motion perception studies -> is it possible to make predictions for other studies?
%What would we predict van Pelt would do for instance?

%%%%%%%%%%%%%%
% Conclusion

### 2015-04-07 src/paper4/paper4.tex
Updated references and introduction
Updated discussion, paper 4

\small {\bf Abstract}
Previously, we have shown that the eye movements that typically accompany self-motion are used to augment self-motion perception. As fixation depth modulates the amplitude of these eye movements, it should be taken into account for a veridical oculomotor estimate of self-motion. To investigate whether the brain compensates for fixation depth, we had  participants ($n = 8$) judge self-motion during different eye movement conditions in the absence of full-field optic flow.  In a 2-AFC task, participants indicated whether the second of two successive passive lateral whole-body translations was longer or shorter than the first. At the start of each movement participants fixated either a nearby or a far away target, which was either body stationary or world stationary during the translation interval. Results show that the perceived translations were shorter for nearby world-fixed gaze compared to far away world-fixed gaze, indicating that eye movements are not properly scaled in self-motion perception. Together with the observation that self-motion perception is not affected by the depth of a body stationary fixation target, we conclude that eye movements are merely a rudimentary cue to self-motion, with a compensation for fixation depth that is partial at best.

@@

In the present study, we investigate how fixation distance influences perception of self-motion during passive lateral translation. Using a psychophysical approach, participants had to indicate whether the second body displacement of two one-second translation intervals was smaller or longer than the first. We show that translation amplitude is perceived smaller when fixating a far compared to a nearby world-centered target, indicating that eye movements are not properly scaled in self-motion perception. Together with the observation that self-motion perception is not affected by the depth of a body-centered fixation target, we conclude that eye movements are merely a rudimentary cue to self-motion, with a compensation for fixation depth that is partial at best.

@@

To investigate this question, \figref{p4:fig3} illustrates psychophysical data on self-motion perception of a single participant for the two-fixation depths in both the world (\figref{p4:fig3}A) and body condition (\figref{p4:fig3}B). Lighter and darker colors indicate which fixation depth was the reference translation (see figure legend). The influence of fixation depth is characterized by a shift of the psychometric functions relative to the 10 \si{\centi\metre} reference translation (i.e. the PSE). For example, the rightward shift of the pink curve in \figref{p4:fig3}A, representing the world-condition, means that with a far target a longer translation (\about15 \si{\centi\metre}) was required to be perceived equivalent to a 10 \si{\centi\metre} reference translation with nearby fixation. Likewise, the leftward shift of the purple curve indicates that a shorter translation with near fixation (\about6 \si{\centi\metre}) is required to be perceived the same as the 10 \si{\centi\metre} reference translation with far fixation. Together, these opposite biases suggest that translations are perceived shorter for fixations further away. For the body condition, no shift of the psychometric curves is visible (\figref{p4:fig3B}), indicating that fixation depth (i.e. near versus far) has no effect in absence of eye movements.

@@

By examining the ratio of these weighting parameters, we remove any depth-independent contributions. In absence of  depth scaling, i.e. when $d_{50} = d_{200}$, the ratio should be one. For perfect compensation, that is when $d_{50} = 50  \wedge d_{200} = 200$, the ratio should be $4$. The actual ratio between $d_{50}$ and $d_{200}$ is plotted for each participant in \figref{p4:fig5}. While two participants show moderate compensation for depth, the majority of participants show no sign of scaling of eye movements by fixation distance. This is consistent with the observation that translations are perceived shorter with far compared to near fixations in the world condition.

@@

% Model results
To quantify the relative depth-dependent scaling of eye movements for nearby and far away fixation targets, we fitted a straightforward linear model to the perceptual responses based on the oculomotor behavior across four conditions. While two participants show partial scaling, the other six participants did not show any sign of scaling. Thus, we conclude that while oculomotor signals provide a robust cue to translation perception, they are not properly scaled by fixation depth.

@@

In our previous experiment we compared translation perception with body-fixed versus world-fixed fixations at near depth (50 \si{\centi\metre}) only \cite{clemens2015a}. \figref{p4:fig6} shows how well our $\alpha_{50}$ parameter explains the data in our previous paper. The positive correlation between the actual PSEs in the previous study and those predicted using the model of the present paper ($\rho = 0.60$, $p = 0.06$) adds confidence to the parameter values presented here. The average difference between the values found here and those reported previously (see \tabref{p4:tab2}) is 12 \textpm 8 percent-points, which is relatively small given the indpendent measurements.

@@

It is important to point out that while the vestibular signal and thus noise is constant for a given translation distance, the noise in the associated oculomotor estimate might change with fixation distance because the magnitude of the eye movement is modulated by fixation depth and may show signal-dependent noise.

>>>>

\small {\bf Abstract}
It has been shown that the compensatory eye movements, that minimize retinal slip during self-motion, also serve as a cue for translation perception. However, to provide an adequate translation estimate, the brain must internally scale these ensuing eye movements by fixation distance. Using a 2-AFC approach, we investigated whether the brain applies this scaling.  Participants ($n = 8$) were translated sideways in the absence of full-field optic flow but with gaze maintained on either a nearby or faraway target, that was either fixed in the world (world-fixed) or moved along with the body (body-fixed). Results show that translations were perceived shorter with gaze on nearby than faraway world-fixed targets, indicating that eye movements are not properly scaled in translation perception. Translation perception was not affected by the depth of body-fixed targets. Taken together, our results suggest that eye movements are merely a rudimentary cue to self-motion, with a compensation for fixation depth that is partial at best.

@@

In the present study, we investigate how fixation distance influences perception of self-motion during passive side-to-side translations. Using a psychophysical approach, participants had to indicate whether the second body displacement of two one-second translation intervals was smaller or longer than the first. We show that translation amplitude is perceived smaller when fixating a far compared to a nearby world-centered target, indicating that eye movements are not properly scaled in self-motion perception. Together with the observation that self-motion perception is not affected by the depth of a body-centered fixation target, we conclude that eye movements are merely a rudimentary cue to self-motion, with a compensation for fixation depth that is partial at best.

@@

\figref{p4:fig3} illustrates psychophysical data on self-motion perception of a single participant for the two-fixation depths in both the world (\figref{p4:fig3}A) and body condition (\figref{p4:fig3}B). Lighter and darker colors indicate which fixation depth was the reference translation (see figure legend). The influence of fixation depth is characterized by a shift of the psychometric functions relative to the 10 \si{\centi\metre} reference translation (i.e. the PSE). For example, the rightward shift of the pink curve in \figref{p4:fig3}A, representing the world-condition, means that with a far target a longer translation (\about15 \si{\centi\metre}) was required to be perceived equivalent to a 10 \si{\centi\metre} reference translation with nearby fixation. Likewise, the leftward shift of the purple curve indicates that a shorter translation with near fixation (\about6 \si{\centi\metre}) is required to be perceived the same as the 10 \si{\centi\metre} reference translation with far fixation. Together, these opposite biases suggest that translations are perceived shorter for fixations further away. For the body condition, no shift of the psychometric curves is visible (\figref{p4:fig3}B), indicating that fixation depth (i.e. near versus far) has no effect in absence of eye movements.

@@

By examining the ratio of these weighting parameters, we remove any depth-independent contributions. In absence of  depth scaling, i.e. when $d_{50} = d_{200}$, the ratio should be one. For perfect compensation, that is when $d_{50} = 50  \wedge d_{200} = 200$, the ratio should be $4$. The actual ratio between $d_{50}$ and $d_{200}$ is plotted for each participant in \figref{p4:fig5}. While two participants show moderate compensation for depth, the majority of participants show no clear sign of scaling of eye movements by fixation distance. This is consistent with the observation that translations are perceived shorter with far compared to near fixations in the world condition.

@@

% Model results
To quantify the relative depth-dependent scaling of eye movements for nearby and far away fixation targets, we fitted a straightforward linear model to the perceptual responses based on the oculomotor behavior across four conditions. While two participants show partial scaling, the other six participants did not show any sign of scaling. Thus, we conclude that oculomotor signals provide a robust cue to translation perception, but are not properly scaled by fixation depth.

@@

In our previous experiment we compared translation perception with body-fixed versus world-fixed fixations at near depth (50 \si{\centi\metre}) only \cite{clemens2015a}. \figref{p4:fig6} shows how well our $\alpha_{50}$ parameter explains the data in our previous paper. The positive correlation between the actual PSEs in the previous study and those predicted using the model of the present paper ($\rho = 0.60$, $p = 0.06$) adds confidence to the parameter values presented here. The average difference between the values found here and those reported previously (see \tabref{p4:tab2}) is 12 \textpm 8 percent-points, which is relatively small given the independent measurements.

@@

It is important to point out that while the vestibular signal and thus its noise level is constant for a given translation distance, the noise in the associated oculomotor estimate might change with fixation distance because the magnitude of the eye movement is modulated by fixation depth and may show signal-dependent noise.
